Empowering the use of preprints in ecology and evolution: an exploration of the first 1000 *EcoEvoRxiv* preprints

### 2023-12-30 ms/ms.qmd
Small edits to sentence structure

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. While preprint use in biology has had a slower uptake [@Berg2016], in ecology and evolution, new discipline specific preprint servers, such as *EcoEvoRxiv*, provide a means by which ecologists and evolutionary biologists can disseminate research findings. Such practices promote open science (e.g., ensuring papers are open access) and facilitate the rapid sharing of new discoveries that can have significant impacts on fundamental and applied knowledge globally (e.g., conservation outcomes).

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [e.g., @Proulx2013]. Importantly, *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives, including accepting multilingual preprints, recognising and accepting registered reports and non-traditional research reports, and promoting peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).

>>>>

Publishing preprints -- papers communicating non-peer-reviewed research findings -- is now an entrenched practice across a multitude of scientific disciplines [@Ginsparg2011]. While preprint use in biology has had a slower uptake [@Berg2016], in ecology and evolution, new discipline specific preprint servers, such as *EcoEvoRxiv*, provide a means by which ecologists and evolutionary biologists can disseminate research findings. Such practices promote open science (e.g., ensuring papers are open access) and facilitate the rapid sharing of new discoveries that can have significant impacts on fundamental and applied knowledge globally.

Preprints attempt to break down barriers to scientific publishing by: 1) increasing the speed, visibility and output of research findings; 2) encouraging constructive and critical feedback from a research community; 3) helping establish precedence of research findings; and 4) removing financial barriers to publication [e.g., @Proulx2013]. *EcoEvoRxiv* is a community driven preprint server that has paved the way for new initiatives, including accepting multilingual preprints, recognising and accepting registered reports and non-traditional research reports, and promoting peer review and community discussion to improve the quality of preprints and speed up their peer-reviewed publication (e.g., Peer Community In -- PCI).
